---
title: "Hormone Project"
author: "Zaynab Mousavian"
date: "Oct. 2022"
output:
  pdf_document: default
  html_document:
    theme:
      bg: white
      fg: black
      base_font:
        google: Prompt
subtitle: Co-expression Network Analysis
---


```{r setup, include=TRUE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

# 1. Loading all data and packages  

```{r setup, include=TRUE,message=FALSE,warning=FALSE}
library(WGCNA)
library(flashClust)
library(ggplot2)
library(reshape)
library(ggrepel)
library(ggpubr)
library(dplyr)
library(readxl)

corMeasure <- "bicor"
networkType <- "signed hybrid"
inputDirectory <- paste0("data/data-June/")
resultsDirectory <- paste0("Results/data-June-V3/Results_Oct/",corMeasure,"-",networkType,"onlyHC")
options(stringsAsFactors = FALSE)

dat_Expr <- read.csv(paste0(inputDirectory,"counts-norm-v3.csv"))
dat_Traits <- read_excel(paste0(inputDirectory,"clinicalData_V3.xlsx"))
dat_Regression <- read.csv(paste0(inputDirectory,"edgeR-model-v3.csv"))
proteinGenes <- dat_Regression[which(dat_Regression$gene_biotype=="protein_coding"),"ENSG"]
dat_Expr <- dat_Expr[which(dat_Expr$gene%in%proteinGenes),]

dat_Traits <- as.data.frame(dat_Traits)
names(dat_Traits) <- dat_Traits[1,]
dat_Traits <- dat_Traits[-1,]
patientIDs <- dat_Traits$PatID
dat_Traits <- dat_Traits[,c("Plasma_S_Estradiol_pg_mL_v3","Plasma_S_Prog_ng_mL_v3")]
names(dat_Traits) <- c("Estradiol","Progesterone")
dat_Traits<- data.frame(apply(dat_Traits, 2, function(x) as.numeric(as.character(x))))
row.names(dat_Traits) <- patientIDs
row.names(dat_Expr) <- dat_Expr$gene
dat_Expr <- dat_Expr[,-1]
names(dat_Expr) <- gsub("X","",names(dat_Expr))

dat_Expr$mad = apply(dat_Expr,1,function(x) mad(x))
sortedGenes = dat_Expr$mad
names(sortedGenes) = row.names(dat_Expr)
sortedGenes = sort(sortedGenes,decreasing = TRUE)
selectedGenes = names(sortedGenes[1:5000])
dat_Expr = dat_Expr[selectedGenes,]
dat_Expr = dat_Expr[,-ncol(dat_Expr)]

head(dat_Expr)
head(dat_Traits)

boxplot(dat_Expr)
```

# 2. Sample clustering 

```{r}
datExpr0 = as.data.frame(t(dat_Expr))
names(datExpr0) = row.names(dat_Expr)
rownames(datExpr0) = colnames(dat_Expr)

gsg = goodSamplesGenes(datExpr0, verbose = 0);
gsg$allOK

if (!gsg$allOK)
{
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], 
                                              collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples],
                                              collapse = ", ")));
  # Remove the offending genes and samples from the data:
  datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}

nHC=nrow(datExpr0)
row.names(datExpr0)=c(paste0("HC0",1:9),paste0("HC",10:nHC))
row.names(dat_Traits)=c(paste0("HC0",1:9),paste0("HC",10:nHC))

sampleTree = flashClust(dist(datExpr0), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
# The user should change the dimensions if the window is too large or too small.

pdf(file = paste0(resultsDirectory,"SampleClustering.pdf"), width = 15, height = 9);
par(cex = 2);
par(mar = c(1,6,2,0))

traitColors=c(rep("#E48725",nHC))
plotDendroAndColors(sampleTree,traitColors,groupLabels="",main = "Sample clustering to detect outliers", sub="", xlab="", cex.main=2, cex.dendroLabels = 1.5,cex.rowText = 1,face=1)

# Plot a line to show the cut
abline(h = 120, col = "red");

# Determine cluster under the line
clust = cutreeStatic(sampleTree, cutHeight = 120, minSize = 10)
table(clust)
# clust 1 contains the samples we want to keep.
keepSamples = (clust==1)
datExpr = datExpr0[keepSamples, ]
datTraits = dat_Traits[keepSamples,,drop=FALSE]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)

# Re-cluster samples
sampleTree2 = flashClust(dist(datExpr), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(datTraits, signed = FALSE);

# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTraits),
                    main = "Sample dendrogram and trait heatmap",cex.main=2)
abline(h = 150, col = "red");
dev.off()
save(datExpr,datTraits,file=paste0(resultsDirectory,"dataInput.RData"))
```

# 3.Choosing the soft-thresholding power 

```{r}
load(paste0(resultsDirectory,"dataInput.RData"))
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))

# Call the network topology analysis function
pearsonFunc<- function(){
  sft <<- pickSoftThreshold(datExpr, powerVector = powers, verbose = 0,
                            networkType=networkType)
}
bicorFunc<- function(){
  sft <<- pickSoftThreshold(datExpr, powerVector = powers, verbose = 0, corFnc="bicor",
                            networkType=networkType, corOptions = list(maxPOutliers =0.1))
}
ifelse(corMeasure=="bicor",bicorFunc(),pearsonFunc())

# Plot the results:
pdf(file=paste0(resultsDirectory,"ScaleFreeTopology.pdf"),width=9,height=5)

par(mfrow = c(1,2));
cex1 = 0.5;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"),cex.main=0.7,cex.lab=1,cex.axis=0.7);
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"),cex.main=0.7,cex.lab=1,cex.axis=0.7)
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=1,col="red")
dev.off()
```

# 4.Co-expression network construction and clustering the network into modules

```{r}
softPower = sft$powerEstimate;
softPower <- ifelse(is.na(softPower),4,softPower)
softPower = 6
pearsonAdjancecy <- function(){
  adjacencyMat <<- adjacency(datExpr, power = softPower, type=networkType);
}
bicorAdjancecy <- function(){
adjacencyMat <<- adjacency(datExpr, power = softPower, corFnc="bicor", type=networkType
                           ,corOptions="maxPOutliers = 0.1");
}
ifelse(corMeasure=="bicor",bicorAdjancecy(),pearsonAdjancecy())

# Turn adjacency into topological overlap
TOM = TOMsimilarity(adjacencyMat,TOMType="signed");
dissTOM = 1-TOM

# Call the hierarchical clustering function
geneTree = flashClust(as.dist(dissTOM), method = "average");
# Plot the resulting clustering tree (dendrogram)
pdf(file=paste0(resultsDirectory,"GeneClustering.pdf"),width=12,height=9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04);

# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 40;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize);
table(dynamicMods)
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
#sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
dev.off()
```

# 5. Merging of close modules

```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = flashClust(as.dist(MEDiss), method = "average");
# Plot the result
pdf(file=paste0(resultsDirectory,"modulesClustering.pdf"),width=7,height=6)

plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
MEDissThres = 0.25
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")
# Call an automatic merging function
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 0)
# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;


MEList = moduleEigengenes(datExpr, colors = mergedColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = flashClust(as.dist(MEDiss), method = "average");
# Plot the result
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

sizeGrWindow(12, 9)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs;
dev.off()
```

# 6. Quantifying module-trait associations

```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
moduleTraitFDR=p.adjust(moduleTraitPvalue,method="fdr")
pdf(file=paste0(resultsDirectory,"ModuleTraits.pdf"),width=8,height=2.7)

# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitFDR, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));

m1=t(moduleTraitCor)
m2=t(textMatrix)
p=labeledHeatmap(Matrix = m1,
               xLabels = names(MEs),
               yLabels = names(datTraits),
               xSymbols = names(MEs),
               colorLabels = FALSE,
               colors = greenWhiteRed(50),
               textMatrix = m2,
               setStdMargins = FALSE,
               cex.text = 0.6,
               cex.main = 0.7,
               cex.axis=0.7,
               cex.lab=0.5,
               cex.lab.x = 0.7,
               cex.lab.y = 0.7,
               zlim = c(-1,1),
               legend.text.direction=45)
dev.off()

# Define variable group containing the group column of datTrait
# Group = as.data.frame(datTraits$Estradiol);
# names(Group) = "Group"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr, datTraits, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(datTraits), sep="");
names(GSPvalue) = paste("p.GS.", names(datTraits), sep="");
probes = names(datExpr)
geneInfo0 = data.frame(Proteins = probes,
                       moduleColor = moduleColors,
                       geneTraitSignificance,
                       GSPvalue)


# Define variable group containing the group column of datTrait
Group = as.data.frame(datTraits$Estradiol);
names(Group) = "Group"
modOrder = order(-abs(cor(MEs, Group, use = "p")));
for (mod in 1:ncol(geneModuleMembership))
{
  oldNames = names(geneInfo0)
  geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]],
                         MMPvalue[, modOrder[mod]]);
  names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
                       paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.Estradiol));
geneInfo = geneInfo0[geneOrder, ]
write.csv(geneInfo, file = paste0(resultsDirectory,"geneInfo.csv"))
save(geneInfo,file=paste0(resultsDirectory,"geneInfo.RData"))

geneInfoMerged=merge(x=geneInfo,y=dat_Regression,by.x="Proteins",by.y="ENSG")
geneInfoMerged = geneInfoMerged[,c("hgnc_symbol",names(geneInfoMerged))]
write.csv(geneInfoMerged,paste0(resultsDirectory,"geneInfo-mergedinfo.csv"))

filterGenes = function(x)
{
  module = x
  columnMM = paste0("MM.",module)
  columnPMM = paste0("p.MM.",module)
  geneInfo_Module = geneInfo[which(geneInfo$moduleColor==module),]
  selectedGenes = intersect(which(geneInfo_Module[,columnMM]>0),
                            which(geneInfo_Module[,columnPMM]<0.05))
  row.names(geneInfo_Module[selectedGenes,])
}
modules = lapply(modNames,function(x) filterGenes(x))
save(modules,modNames,file=paste0(resultsDirectory,"modules.RData"))
```

# 7. Extracting information about modules of interest

```{r}
l=load(file=paste0(resultsDirectory,"geneInfo.RData"))
load(file=paste0(resultsDirectory,"modules.RData"))
module="blue"
column = match(module, modNames);
moduleGenes = modules [[column]] #94 ensemble gene ids

geneInfoModuleMerged=merge(x=geneInfo[moduleGenes,],y=dat_Regression,by.x="Proteins",by.y="ENSG")
geneInfoModuleMerged = geneInfoModuleMerged[,c("hgnc_symbol",names(geneInfoModuleMerged))]
write.csv(geneInfoModuleMerged,paste0(resultsDirectory,"-",module,"-mergedinfo.csv"))

gene.symbols <- dat_Regression[match(moduleGenes,dat_Regression$ENSG),"hgnc_symbol"]
write.csv(gene.symbols,file=paste0(resultsDirectory,module,".csv"))
geneInfoModule <- geneInfo[moduleGenes,]
geneInfoModule$symbol <- gene.symbols
blueModule <- geneInfoModule$symbol

write.csv(geneInfoModule,paste0(resultsDirectory,"-",module,"-info.csv"))

geneInfoModule_filtered <- geneInfoModule %>% filter(p.MM.blue<0.05&MM.blue>0.5)
blueModule_filtered <- geneInfoModule_filtered$symbol

geneInfoModuleMerged_filtered=merge(x=geneInfoModule_filtered[moduleGenes,],y=dat_Regression,by.x="Proteins",by.y="ENSG")
geneInfoModuleMerged_filtered = geneInfoModuleMerged_filtered[,c("hgnc_symbol",names(geneInfoModuleMerged_filtered))]

write.csv(geneInfoModuleMerged_filtered,paste0(resultsDirectory,"-",module,"-filtered-mergedinfo.csv"))

write.csv(geneInfoModule_filtered,paste0(resultsDirectory,"-",module,"-filtered-info.csv"))

module="green"
column = match(module, modNames);
moduleGenes = modules [[column]] #94 ensemble gene ids

geneInfoModuleMerged=merge(x=geneInfo[moduleGenes,],y=dat_Regression,by.x="Proteins",by.y="ENSG")
geneInfoModuleMerged = geneInfoModuleMerged[,c("hgnc_symbol",names(geneInfoModuleMerged))]
write.csv(geneInfoModuleMerged,paste0(resultsDirectory,"-",module,"-mergedinfo.csv"))

gene.symbols <- dat_Regression[match(moduleGenes,dat_Regression$ENSG),"hgnc_symbol"]
write.csv(gene.symbols,file=paste0(resultsDirectory,module,".csv"))
geneInfoModule <- geneInfo[moduleGenes,]
geneInfoModule$symbol <- gene.symbols
greenModule <- geneInfoModule$symbol

write.csv(geneInfoModule,paste0(resultsDirectory,"-",module,"-info.csv"))

geneInfoModule_filtered <- geneInfoModule %>% filter(p.MM.green<0.05&MM.green>0.5)
greenModule_filtered <- geneInfoModule_filtered$symbol

geneInfoModuleMerged_filtered=merge(x=geneInfoModule_filtered[moduleGenes,],y=dat_Regression,by.x="Proteins",by.y="ENSG")
geneInfoModuleMerged_filtered = geneInfoModuleMerged_filtered[,c("hgnc_symbol",names(geneInfoModuleMerged_filtered))]
write.csv(geneInfoModuleMerged_filtered,paste0(resultsDirectory,"-",module,"-filtered-mergedinfo.csv"))

write.csv(geneInfoModule_filtered,paste0(resultsDirectory,"-",module,"-filtered-info.csv"))
```

# 8. Finding overlap between modules of interest and results of regression analysis

```{r}
library(dplyr)
table=data.frame()

regressionResults <- read.csv(paste0(inputDirectory,"edgeR-model-v3.csv"))
FDR_significant_down <- filter(regressionResults ,logFC_E2<0 & FDR_E2<0.05)
Pval_significant_down <- filter(regressionResults ,logFC_E2<0 & PValue_E2<0.05)
all_down <- filter(regressionResults ,logFC_E2<0)

module <- read.csv(paste0(resultsDirectory,"-blue-info.csv"))
b3=length(module[which(module$Proteins%in%FDR_significant_down$ENSG),"symbol"])
b2=length(module[which(module$Proteins%in%Pval_significant_down$ENSG),"symbol"])
b1=length(module[which(module$Proteins%in%all_down$ENSG),"symbol"])
a3=dim(FDR_significant_down)[1]
a2=dim(Pval_significant_down)[1]
a1=dim(all_down)[1]
a0=dim(module)[1]

table=rbind(table,c(a0,a1,a2,a3,b1,b2,b3))

module <- read.csv(paste0(resultsDirectory,"-blue-filtered-info.csv"))
b3=length(module[which(module$Proteins%in%FDR_significant_down$ENSG),"symbol"])
b2=length(module[which(module$Proteins%in%Pval_significant_down$ENSG),"symbol"])
b1=length(module[which(module$Proteins%in%all_down$ENSG),"symbol"])
a3=dim(FDR_significant_down)[1]
a2=dim(Pval_significant_down)[1]
a1=dim(all_down)[1]
a0=dim(module)[1]

table=rbind(table,c(a0,a1,a2,a3,b1,b2,b3))

blueModule = FDR_significant_down$hgnc_symbol

library(dplyr)
regressionResults <- read.csv(paste0(inputDirectory,"edgeR-model-v3.csv"))
FDR_significant_up <- filter(regressionResults ,logFC_E2>0 & FDR_E2<0.05)
Pval_significant_up <- filter(regressionResults ,logFC_E2>0 & PValue_E2<0.05)
all_up <- filter(regressionResults ,logFC_E2>0)

module <- read.csv(paste0(resultsDirectory,"-green-info.csv"))
b3=length(module[which(module$Proteins%in%FDR_significant_up$ENSG),"symbol"])
b2=length(module[which(module$Proteins%in%Pval_significant_up$ENSG),"symbol"])
b1=length(module[which(module$Proteins%in%all_up$ENSG),"symbol"])
a3=dim(FDR_significant_up)[1]
a2=dim(Pval_significant_up)[1]
a1=dim(all_up)[1]
a0=dim(module)[1]

table=rbind(table,c(a0,a1,a2,a3,b1,b2,b3))

module <- read.csv(paste0(resultsDirectory,"-green-filtered-info.csv"))
b3=length(module[which(module$Proteins%in%FDR_significant_up$ENSG),"symbol"])
b2=length(module[which(module$Proteins%in%Pval_significant_up$ENSG),"symbol"])
b1=length(module[which(module$Proteins%in%all_up$ENSG),"symbol"])
a3=dim(FDR_significant_up)[1]
a2=dim(Pval_significant_up)[1]
a1=dim(all_up)[1]
a0=dim(module)[1]

table=rbind(table,c(a0,a1,a2,a3,b1,b2,b3))
names(table)= c("Number of genes in module", "number of all negative logFC from Olga", "number of sig Pvalue from Olga", "number of sig FDR from Olga", "overlap between logFC and module", "overlap between Pvalue and module", "overlap between FDR and module")
row.names(table)=c("blue","blue-filtered","green","green-filtered")
write.csv(table,file=paste(resultsDirectory,"-details about number of genes.csv"))
greenModule = FDR_significant_up$hgnc_symbol
```

# 9. Functional enrichment analysis for modules of interest

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichR)
library(cowplot)
library(xlsx)
gene_list = blueModule_filtered
col <- "blue"
thr <- 0.05
threshold <- 1.3
point_size <- 6
text_size <-
dbs <- listEnrichrDbs()
head(dbs)
BPs <- c("GO_Biological_Process_2023")
pathways <- c("KEGG_2021_Human")
databases <- c("DisGeNET")
TFs <- c("TRRUST_Transcription_Factors_2019")
species <- "Homo sapiens"

enriched <- enrichr(gene_list, BPs)
result <- enriched[[1]]
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "GO terms-Blue module",append = T,row.names = F)

result$val <- -log10(result$Adjusted.P.value)
result <- result[1:10,]
result$Term <- unlist(lapply(result$Term, function(x) sub("\\s*\\(.*", "", x)))
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))   # This trick update the factor levels
p1 = ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5,color="white") +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("")

enriched <- enrichr(gene_list, pathways)
result <- enriched[[1]]
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "KEGG pathways-Blue module",append = T,row.names = F)

result$val <- -log10(result$Adjusted.P.value)
result <- result[1:5,]
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))  # This trick update the factor levels

p2 <-  ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5,color="white") +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("")

enriched <- enrichr(gene_list, TFs)
result <- enriched[[1]]
result <- result %>% filter(grepl("human",Term))
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "TRRUST-Blue module",append = T,row.names = F)
result$val <- -log10(result$Adjusted.P.value)
result <- result[1:5,]
result$Term <- gsub("human","",result$Term)
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))    # This trick update the factor levels

p3 <- ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5,color="white") +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("")

gene_list = greenModule_filtered
col <- "green"
dbs <- listEnrichrDbs()
head(dbs)
BPs <- c("GO_Biological_Process_2021")
pathways <- c("KEGG_2021_Human")
databases <- c("DisGeNET")
TFs <- c("TRRUST_Transcription_Factors_2019")
species <- "Homo sapiens"

enriched <- enrichr(gene_list, BPs)
result <- enriched[[1]]
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "GO terms-Green module",append = T,row.names = F)

result$val <- -log10(result$Adjusted.P.value)
result <- result[1:10,]
result$Term <- unlist(lapply(result$Term, function(x) sub("\\s*\\(.*", "", x)))
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))   # This trick update the factor levels

p4 = ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5) +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("GO biological process")


enriched <- enrichr(gene_list, pathways)
result <- enriched[[1]]
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "KEGG pathways-Green module",append = T,row.names = F)

result$val <- -log10(result$Adjusted.P.value)
result <- result[1:5,]
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))  # This trick update the factor levels

p5 <-  ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5) +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("KEGG Pathways")

enriched <- enrichr(gene_list, TFs)
result <- enriched[[1]]
result <- result %>% filter(grepl("human",Term))
write.xlsx(result,file=paste0(resultsDirectory,"Functional enrichment results.xlsx"),sheetName = "TRRUST-Green module",append = T,row.names = F)
result$val <- -log10(result$Adjusted.P.value)
result <- result[1:5,]
result$Term <- gsub("human","",result$Term)
result$num <- unlist(lapply(result$Overlap, function(x) unlist(strsplit(x, "/"))[1]))

result <- result %>%
  arrange(val) %>%    # First sort by val. This sort the dataframe but NOT the factor levels
  mutate(Term=factor(Term, levels=Term))    # This trick update the factor levels

p6 <- ggplot(result, aes(x=Term, y=val))+
    geom_segment( aes(xend=Term, yend=0)) +
    geom_point(size=point_size, color=col) +
    geom_text(aes(label = num), size=3, vjust = 0,hjust= 0.5) +
    geom_hline(yintercept = threshold, color = "red", linetype = "dashed", size = 1) +
    coord_flip() +
    theme_bw() +
    ylab("-log10(FDR pValue)")+
    xlab("TRRUST")


pdf(paste0(resultsDirectory,"Functional enrichment results-ModulesFiltered.pdf"),width=12,height = 6)
plot_grid(p4,p1,p5,p2,p6,p3,nrow=3,ncol=2,align="hv",rel_heights = c(1,0.5,0.5),rel_widths = c(1,1))
dev.off()
```

# 10. Preservation analysis 

```{r}
l=load(paste0(resultsDirectory,"dataInput.RData"))
datExprFirst=datExpr

l=load("results/data-June-V2/Final results/bicor-signed hybridonlyHCdataInput.RData")
datExprSecond=datExpr

datExprFirst=datExprFirst[,intersect(names(datExprFirst),names(datExprSecond))]
datExprSecond=datExprSecond[,intersect(names(datExprFirst),names(datExprSecond))]

file=read.csv(paste0(resultsDirectory,"geneInfo.csv"))
#file=read.csv("GeneInfo_Limma_Ordering(k_logFC).csv")

row.names(file)=file$Proteins
FirstColors=unlist(lapply(names(datExprFirst),function(x) file[x,"moduleColor"]))

setLabels = c("First", "Second");
multiExpr = list(First = list(data = datExprFirst), Second = list(data = datExprSecond));
multiColor = list(First = FirstColors)

system.time( {
  mp = modulePreservation( multiExpr, multiColor,
                          referenceNetworks = 1,
                          nPermutations = 5,
                          randomSeed = 1,
                          quickCor = 0,
                          verbose = 3)
} );


ref=1; test = 2
Obs.PreservationStats= mp$preservation$observed[[ref]][[test]]
Z.PreservationStats=mp$preservation$Z[[ref]][[test]]

modColors = rownames(Obs.PreservationStats)
moduleSize = Obs.PreservationStats$moduleSize
# we will omit the grey module (background genes)
# and the gold module (random sample of genes)
selectModules = !(modColors %in% c("grey", "gold"))
# Text labels for points
point.label = modColors[selectModules]

#Composite preservation statistics
medianRank=Obs.PreservationStats$medianRank.pres
Zsummary=Z.PreservationStats$Zsummary.pres
dataPlot = data.frame(size=moduleSize[selectModules],Zsummary=Zsummary[selectModules],color=modColors[selectModules])


pdf(paste0(resultsDirectory,"preservation.pdf"),width=4,height = 2.4)
ggplot(dataPlot,aes(x=size,y=Zsummary,color=color))+
  geom_point(size=3)+
  scale_color_identity()+
  scale_y_continuous(trans = "log10")+
  labs(x="Module Size (genes)")+
  geom_hline(yintercept = 10,color="red",linetype="dashed")+
  theme_classic()+
  theme(panel.border = element_rect(colour = "black",fill="NA"),axis.title=element_text(size=8),panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
dev.off()
```
```

