---
title: "#6620"
format: html
toc: true
toc-expand: 3
editor_options: 
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"    
# bibliography: references.bib 
---

```{r}
#| code-fold: true
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(DT)
```

## Support request

Delineating molecular mechanisms of how female sex hormones influence the integrity of the female genital mucosa.

::: {.callout-note collapse="true"}
## More information: Support request

In a previous study, we studied how the use of injectable hormonal contraceptive depo-medroxyprogesterone (DMPA) affects the transcriptome of the female genital tract (Bradley et al, PLoS Pathogens 2022). For this, ectocervical tissue biopsies were collected from 96 Kenyan women and analyzed by RNA-sequencing. Interestingly, the transcriptomics data indicate that endogenous serum levels of the female sex hormones progesterone and estradiol influence the transcriptome. This was briefly touched upon in the published paper mentioned above, but we now wish to further explore these findings and correlate gene expression to serum hormone levels. We have access to RNA-sequencing data from 96 tissue samples, and an additional 110 biopsies from the same study subjects but collected at a different time point are now being analyzed using RNA-sequencing. We expect this additional data to arrive within the next few weeks. In addition, the serum hormone levels for estradiol and progesterone are available for the majority of samples. We would like assistance in data handling and processing, including statistical help.

Full support description available via [Redmine](https://projects.nbis.se/issues/6620)

-   Request by: Frideborg Bradley \<[frideborg.bradley\@ki.se](mailto:frideborg.bradley@ki.se){.email}\>
-   Principal Investigator: Kristina Broliden \<[kristina.broliden\@ki.se](mailto:kristina.broliden@ki.se){.email}\>
-   Organization: KI
-   NBIS staff: Olga Dethlefsen \<[olga.dethlefsen\@nbis.se](mailto:olga.dethlefsen@nbis.se){.email}\>
:::


## Available data

### V2 and V3

Data include measurements from two phases, V2 and V3:

-   V2 luteal phase (n = 58)
-   V3 follicular phase (n = 66)

### Clinical data

Key variables:

-   Hormone levels: estradiol \[pg/ml\] and progesterone \[ng/ml\]
-   Age in years, it is same at both visit, V2 and V3.
-   Bacterial Vaginosis, BV, is a clinical diagnosis for cervicovaginal dysbiosis (simplified explanation). It is diagnosed using Nugent score, and each woman is assigned a number 0-10. 0-3= normal, 4-6= intermediate, 7-10 = BV. Coding: 0 = Normal, 1 = intermediate, 2 = BV. There is a separate BV-diagnosis available for V2 and V3.

::: {.callout-note collapse="true"}
### More information: Plasma hormone analysis

Blood samples were collected by venipuncture using heparin as an anticoagulant, and plasma was isolated using a Ficoll density gradient. For the V2 samples, plasma levels for E2 and P4 were measured using the Milliplex Map Steroid/Thyroid Hormone Magnetic Bead Panel (Millipore, Merck, Darmstadt, Germany). The lower limits of detection for E2 and P4 were 20 pg/ml and 0.09 ng/ml respectively. Values below this were reported as "below Lower Limit of Detection" and were assigned a value of ½ LLD, being 10 pg/ml and 0.05 ng/ml for E2 and P4 respectively.

For the V3 samples, plasma E2 and P4 levels were measured using electrochemiluminescence immunoassays (Roche Diagnostics) at the accredited Karolinska University Laboratory as previously described (Plos Pathogens peket). The LLDs for E2 and P4 were 22 pg/mL and 0.05 ng/mL, respectively. Values \<22 pg/mL (E2) and \<0.05 ng/mL (P4) were reported as "Below LLD", but for statistical reasons they were assigned values of 22 pg/mL and 0.05 ng/mL for E2 and P4, respectively.
:::

### Gene expression

Gene expression counts were provided by BEA facility.

-   BEA gene expression coding: BEA22P129_KB, run 2023-04
-   BEA gene expression coding: BEA19P035_KB, sequenced previously, bioinformatics pipeline re-run 2023-04 to match V2

### Genes of interest

```{r}
#| echo: false
#| message: false
#| warning: false

file <- "../data/box/Selected genes to OD 20230504.csv"
genes_int <- read_csv(file)

datatable(genes_int, caption = "List of genes of interest.")

```

## Methods

- A set of clean data sets were prepared for V2 and V3 based on the data provided by the BEA and the group. This included raw count data matching clinical data, extracted relevant clinical data and genes annotations. For the latter, `biomaRt` package was used to annotate Ensembl gene IDs with chromosome name, start and end position, strand, symbol, description and gene biotype as well as corresponding Entrez Gene ID. 
- `edgeR` was used for analysis of the gene expression data. 
- Lowly expressed genes were filtered out retaining genes with a minimum count of 10 in at least 70% of samples. 
- TMM normalization wass used for estimating relative RNA levels from the count data and normalized counts were $log2$ transformed. 
- The normalized counts were used to examine the samples for outliers and relationships, using PCA. Association between PCA components and covariates was examined using linear regression model for numerical variables and using one-way ANOVA for categorical variables.
- `edgeR` package was further used for association analysis between genes and E2 and P4 levels, adjusting for BV. Quasi-likelihood negative binomial generalized log-linear model was fitted to the data and gene-wise statistical test for a given coefficient (E2 or P4) was run to identify significant genes. P-values were adjusted for multiple testing using Benjamin & Hochberg procedure. 


## Results

### Data wrangling

-   [prepare-data-sets.html](prepare-data-sets.html)

### EDA, exploratory data analysis

-   [eda.html](eda.html)

### Statistical modelling

-   [model.html](model.html)
